\section{Shende's argument}
In order to compare the Hodge structures of the De Rham and the Betti moduli spaces, we employ Shende's idea presented in \cite[]{Shende-weights}. Shende replaced a topological space, in particular a complex manifold, with a simplicial scheme and use this simplicial decomposition to give a presentation of the Betti moduli space as a simplicial scheme. The main point of the construction is that, while the Betti moduli space doesn't admit a universal local system, since such a local system would not be of algebraic nature, this simplicial replacement does. As we saw in the previous system, universal bundles are the key to comparing the cohomology rings because the K\"unneth components of their Chern classes are generators.

In what follows, we will first explain in detail the idea behind Shende's construction. Then we will see how it applies to our case, when instead of a complex manifold as studied in \cite[]{Shende-weights} we have a stacky curve.

\subsection{A simplicial interpretation of the stack of local systems}
Let $X$ be a complex manifold, choose a base point $v_0 \in X$ for the fundamental group of $X$. For an algebraic group $G$, the stack of $G$-local systems on $X$ is the quotient
\begin{equation} \label{stack of local systems}
[\Hom(\pi_{1}(X, v_0), G) / G],
\end{equation}
where $\pi_{1}(X, v_0)$ is seen as a discrete algebraic group over $\Spec \mathbb{C}$.

Of course, the definition \eqref{stack of local systems} makes sense for any topological space $X$. By replacing $X$ with a simplicial decomposition, we can replace \eqref{stack of local systems} with an equivalent quotient stack.

Notice that we have a simplicial interpretation of the fundamental group. Indeed, if $\Delta_X$ is a simplicial set and $v_0 \in \Delta_X^0$ is a vertex, define $E(\Delta_X, v_0)$ the \textit{edge-path group} at $v_0$ as the group generated by loops of $1$-simplices pointed at $v_0$ under path-composition, modulo the relation killing boundaries of $2$-simplices
\begin{equation}
% Use \Biggl and \Biggr for larger angled parentheses 
% and \mid for a smaller separator than \middle|
E(\Delta_X, v_0) = \Biggl\langle
    \tikzpath{
      \node[dot, label=below:{\tiny $v_1$}] (v1) at (1,0) {};
      \node[dot, label=below:{\tiny $v_n$}] (vn) at (2.2,0) {};
      \node[dot, label=below:{\tiny $v_0$}] (v0e) at (3.2,0) {};
      \draw[-{Stealth[scale=0.5]}] (v0) -- (v1);
      \draw[dashed] (v1) -- (vn);
      \draw[-{Stealth[scale=0.5]}] (vn) -- (v0e);
    }
  \:\bigg|\:
  \begin{aligned}
    & 
      \tikzpath{
        \node[dot, label=below:{\tiny $v_n$}] (vn) at (1,0) {};
        \node[dot, label=below:{\tiny $v_0$}] (v0m) at (1.8,0) {};
        \node[dot, label=below:{\tiny $v'_1$}] (v1p) at (2.6,0) {};
        \node[dot, label=below:{\tiny $v'_{n'}$}] (vnp) at (3.8,0) {};
        \node[dot, label=below:{\tiny $v_0$}] (v0e) at (4.6,0) {};
        \draw[dashed] (v0) -- (vn);
        \draw[-{Stealth[scale=0.5]}] (vn) -- (v0m);
        \draw[-{Stealth[scale=0.5]}] (v0m) -- (v1p);
        \draw[dashed] (v1p) -- (vnp);
        \draw[-{Stealth[scale=0.5]}] (vnp) -- (v0e);
      }
    = 
      \tikzpath{
        \node[dot, label=below:{\tiny $v'_{n'}$}] (vnp) at (1,0) {};
        \node[dot, label=below:{\tiny $v_0$}] (v0e) at (2,0) {};
        \draw[dashed] (v0) -- (vnp);
        \draw[-{Stealth[scale=0.5]}] (vnp) -- (v0e);
      }
      \cdot
      \tikzpath{
        \node[dot, label=below:{\tiny $v_n$}] (vn) at (1,0) {};
        \node[dot, label=below:{\tiny $v_0$}] (v0e) at (1.8,0) {};
        \draw[dashed] (v0) to (vn);
        \draw[-{Stealth[scale=0.5]}] (vn) -- (v0e);
      }
    , \\
    & \begin{tikzpicture}[baseline={(v0.center)}, xscale=0.6, yscale=1]
      \tikzset{dot/.style={circle, fill, inner sep=0.8pt}}
      \node[dot] (v0) at (0,0) {}; 
      \node[dot] (v1) at (1,0) {}; \node[dot] (v2) at (2,0) {};
      \draw[-{Stealth}] (v0) -- node[below]{\tiny $\partial_1 \sigma$} (v1);
      \draw[-{Stealth}] (v1) -- node[below]{\tiny $\partial_0 \sigma$} (v2);
    \end{tikzpicture}
    =
    \begin{tikzpicture}[baseline={(v0.center)}, xscale=0.6, yscale=1]
      \tikzset{dot/.style={circle, fill, inner sep=0.8pt}}
      \node[dot] (v0) at (0,0) {};
      \node[dot] (v1) at (1,0) {};
      \draw[-{Stealth}] (v0) -- node[below]{\tiny $\partial_2 \sigma$} (v1);
    \end{tikzpicture} \quad
    \forall \sigma \in X_2
  \end{aligned}
\Biggr\rangle.
\end{equation}

\begin{rmk} \label{iso fundamental group - edge-path group}
There is an isomorphism $\pi_1(|\Delta_X|, v_0) \simeq E(\Delta_X, v_0)$.
\end{rmk}

\begin{prop}\label{prop: stack of local systems - simplicial version}
Let X be a path-connected topological space. Suppose that $\pi_{1}(X, v_0)$ is finitely presented, e.g. if $X$ is compact. Let $\Delta_X$ be a simplicial set whose geometric realization is homotopy equivalent to $X$, having $v_0$ as a vertex. Call $X_i$ the $i$-th skeleton of $\Delta_X$.
Then the isomorphism $\pi_1(X, v_0) \simeq E(\Delta_X, v_0)$ induces an equivalence of quotient stacks
\begin{equation}
\left[ \Hom(\pi_1(X, v_0), G) / G \right] 
\xrightarrow[ \cong ]{\quad \mu \quad}
\left[ \left\{ (g_e)_{e \in X_1} \mid g_{\partial_1 \sigma} 
\, g_{\partial_0 \sigma}^{-1} \, g_{\partial_2 \sigma} = 1 
\quad \forall \ \sigma \in X_2 \right\} / G^{\# X_0} \right],
\end{equation}
where $G^{\# X_0}$ acts on $F \coloneqq \left\{ (g_e)_{e \in X_1} \mid g_{\partial_1 \sigma} \, g_{\partial_0 \sigma}^{-1} \, g_{\partial_2 \sigma} = 1 \quad \forall \ \sigma \in X_2 \right\}$ by edge-boundaries, i.e.
\begin{equation}
(g_v)_{v \in X_0} \cdot (g_e)_{e \in X_1} = (g_{\partial_0 e} \, g_e \, g_{\partial_1 e}^{-1})_{e \in X_1}.
\end{equation}
\end{prop}

\begin{rmk}
Both $\Hom(\pi_1(X, v_0), G)$ and $F$ can be expressed as closed subvarieties of some $G^N$, so indeed the ones described in \zcref{prop: stack of local systems - simplicial version} are well defined algebraic stacks.
\end{rmk}

Using \zcref{iso fundamental group - edge-path group}, in proving \zcref{prop: stack of local systems - simplicial version} we replace $\pi_1(X, v_0)$ by $E(\Delta_X, v_0)$.

\begin{proof}
We actually prove that the quotients prestacks are equivalent. Namely, we show that for $S$ scheme over $\mathbb{C}$ the groupoids
\begin{equation}
\begin{tikzcd}
	{G(S) \times \Hom(E(\Delta_X, v_0), G)(S)} & {\Hom(E(\Delta_X, v_0), G)(S)}
	\arrow[shift right, from=1-1, to=1-2]
	\arrow[shift left, from=1-1, to=1-2]
\end{tikzcd}
\end{equation}
and
\begin{equation}
\begin{tikzcd}
	{G(S)^{\# X_0} \times F(S)} & {F(S)}
	\arrow[shift right, from=1-1, to=1-2]
	\arrow[shift left, from=1-1, to=1-2]
\end{tikzcd}
\end{equation}
are equivalent, functorially in $S$.
\end{proof}

